---
title: "edger_use"
author: "Kaitlyn Strickfaden"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{edger_use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<br>

```{r options, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 8,
  fig.height = 6,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = F}
library(edger)
library(dplyr)
library(imager)
library(progressr)
library(qpdf)
library(stringr)
```

<br>

#  Using the *edger* Package

The `edger` package provides a simple method for extracting the outlines of an object of interest in one image and superimposing the same outlines onto new images. This vignette describes the various functions and how to use them. For a detailed description of the methodology behind the functions, refer to the `edger` methodology [vignette](https://github.com/kaitlynstrickfaden/edger/blob/master/vignettes/edger_methodology.md).

<br>

It's really important that the object you want to extract contrasts with the background. If the image is overexposed or underexposed, or if the object is a similar color to the background, the methods described below won't work quite as well.


## Functions

<br>

Right now, this package includes three functions:

- `edger_single`: find and recolor outlines in just one image, with the option to save the output image. Useful for tinkering with threshold values until you find the right one for your image.
- `edger_testr`: an interactive function to extract coordinates for the region(s) of interest (ROI) and find appropriate threshold values. The coordinates and values are output in a format that can be used by `edger_single` and `edger_multi`.
- `edger_multi`: find and recolor outlines in one image, then superimpose those outlines onto a set of other images.

<br>

---

<br>

## edger_single

The `edger_single` function takes the file path to an image as an input, asks the user to either draw or input a region of interest (**ROI**), and outputs a new image with the edges in that ROI recolored. There are a few other arguments to allow for more customization though. These are:

<br>

- `th`: a numeric input setting the threshold value for determining edges. Generally, this is a value greater than 0 and less than 0.20. Lower values capture weaker edges, and higher values capture stronger edges. The default is 0.10.
- `roi`: an argument for supplying the coordinates of ROIs without having to draw them each time. The default is NULL and will cause the function to launch the user interface for drawing a ROI on the image. If the user decides to input the ROI, it must be input as a list containing a data frame, and the data frame must contain the minimum x, minimum y, maximum x, and maximum y coordinate of the ROI. Each ROI must be in its own row of the data frame. It may seem weird that `roi` has to be a list of a single data frame, but this is just so the object type of `roi` is the same whether you have one ROI or several (as is possible with `edger_multi`). 
- `regions`: a numeric input for setting the number of ROIs to draw. The default is 1. If a user has multiple objects in an image but does not want to capture the area between the objects, he or she can increase the value of `regions` to capture each object separately.
- `shift`: a vector of length 2 for the number of pixels the drawn object should be moved left, right, up, or down. The first value in the pair will be the left-right shift, and the second value will be the up-down shift. The default is no shift (`c(0,0)`).
- `rotate`: a numeric indicating how many degrees to rotate the recolored pixels. Negative values rotate the pixels counter-clockwise, and positive values rotate the pixels clockwise. The default is 0.
- `color`: a character string for the desired color of the recolored pixels. The default is red, but that may not be a preferred color in all contexts.
- `save`: a Boolean (T/F) indicating if the recolored image should be saved. The default is False, but if it is set to True, the recolored image will be saved to the working directory with "_edger" appended to the original file name.

<br> 

Let's see the `edger_single` function in action. First we'll load in a file path to a raw image:

<br>

```{r load image 1}
im1 <- "../images/image03.jpg"
par(mar = c(0,0,0,0))
plot(imager::load.image(im1), axes = F)
```

<br>

If we provide this file path to the `edger_single` function, this is the result:

<br>

```{r edger_single dummy 1, echo = T, eval = FALSE}
edger::edger_single(im1)
```

```{r edger_single 1, echo = F}
roi1 <- data.frame(x1 = 894, y1 = 538, x2 = 974, y2 = 1219)
edger_single(im1, roi = list(roi1))
```
 
 <br>

Note also that the `edger_single` function outputs the amount of time it took for the function to run.

<br>

A threshold value of 0.10 looks pretty good for this image, but let's adjust some of the inputs and try again just to see what happens. We'll also set the ROI externally. When `imager` plots an image, it starts by plotting the top-left corner and then working its way right and then *down*, so it's a bit different from normal plotting convention. But if you use `imager`'s `grabRect` function, it will output those coordinates correctly.

<br>

```{r edger_single dummy 2, echo = T, eval = FALSE}
roi1 <- data.frame(x1 = 894, y1 = 538, x2 = 974, y2 = 1219)
edger::edger_single(im1, roi = list(roi1), th = 0.07, color = "cyan1")
```

```{r edger_single 2, echo = F}
edger_single(im1, roi = list(roi1), th = 0.07, color = "cyan1")
rect(roi1[,1], roi1[,2], roi1[,3], roi1[,4], lwd = 2)
text(x = roi1[,1] - 100, roi1[,2] - 40, labels = str_glue("({roi1[,1]}, {roi1[,2]})"))
text(x = roi1[,3] + 120, roi1[,4] + 40, labels = str_glue("({roi1[,3]}, {roi1[,4]})"))
```

<br>

Notice that by decreasing the `th` input, we picked up some weaker edges around the measuring stake. 

<br>

If we wanted to recolor the edges of that second measuring stake too, we can do that by increasing the `regions` input and adding another row to our `roi` data frame.

<br>

```{r edger_single 3, echo = T}
roi2 <- data.frame(x1 = 457, y1 = 501, x2 = 526, y2 = 1105)
edger::edger_single(im1, roi = list(rbind(roi1, roi2)), regions = 2, color = "purple")
```


<br>

---

<br>

## edger_testr

The `edger_testr` function is an extension of the `edger_single` function. Rather than input just a single file path, the user can input a vector containing paths to multiple images to the `imagepaths` argument. The `edger_testr` function launches an interactive interface to the `edger_single` function. It will use the default `th` from the `edger_single` function (0.10) to recolor edges on the first instance; then, the function will ask the user if the `th` used was a good value. If the user inputs "Y", the function will save the ROI coordinates and `th` to a list of data frames. If the user inputs "N" (or any other value), the interface will ask for a new `th` and rerun the `edger_single` function. It will run this same procedure on the rotate and shift inputs (default is no rotation or shift). It will move on to a new image once a threshold, rotate, and shift value have been selected for the current image. Once the function has gotten through all of the images in `imagepaths`, the `edger_testr` function will output a list containing coordinates to ROIs, threshold values, shift values, and rotate values.

<br>

---

<br>

## edger_multi

<br>

The final function, the `edger_multi` function, is the workhorse of the `edger` package. It allows the user to take the edges he or she found in one image and recolor those pixels in a new set of images with any shifting or rotation as desired. It will save the recolored images with "_edger" appended to the original file name, so you won't lose any of your raw images. These are the arguments for the `edger_multi` function:

- `images`: a vector containing file paths to images. The first image is referred to as the "reference" image; this is the image in which edges will be found that will then be drawn on to the rest of the images. If `ref_images` is greater than 1, then the function will use *n* images as the reference images.
- `ref_images`: a numeric input for the number of reference images to be used. The default is one. 
- `roi`: an argument for supplying the coordinates of ROIs without having to draw them each time. The default is NULL and will cause the function to launch the user interface for drawing a ROI on the image. If the user decides to input the ROI, it must be input as a list containing a data frame for each reference image, and each data frame must contain the minimum x, minimum y, maximum x, and maximum y coordinate of the ROI. Each ROI must be in its own row of the data frame.
- `th`: a numeric input setting the threshold value for determining edges. Generally, this is a value greater than 0 and less than 0.20. Lower values capture weaker edges, and higher values capture stronger edges. The default is 0.10.
- `regions`: a numeric input for setting the number of ROIs to draw. The default is 1. If a user has multiple objects in an image but does not want to capture the area between the objects, he or she can increase the value of `regions` to capture each object separately.
- `shift`: a vector of length 2 for the number of pixels the drawn object should be moved left, right, up, or down. The first value in the pair will be the left-right shift, and the second value will be the up-down shift. The default is no shift (`c(0,0)`).
- `rotate`: a numeric indicating how many degrees to rotate the recolored pixels. Negative values rotate the pixels counter-clockwise, and positive values rotate the pixels clockwise. The default is 0.
- `color`: a character string for the desired color of the recolored pixels. The default is red, but that may not be a preferred color in all contexts.
- `show_image`: a Boolean (T/F) indicating if the images should be plotted after being recolored. The default is True. The function takes longer to run if `show_image` is True, but plotting images allows the user to diagnose issues with the recoloring.
- `process`: a character input of either "sequential" or "parallel" indicating if the images should be processed on one core or many. The number of cores used is determined by the "cores" argument. The default is "sequential".
- `cores`: a numeric indicating how many cores on which to process images. The default is 1 (i.e. sequential).

<br>

It looks like a lot of inputs, but most of them are the same as the arguments for `edger_single`. 

<br>

Let's load in a few more images:

<br>

```{r load images 2 and 3}
im2 <- "../images/image04.jpg"
im3 <- "../images/image05.jpg"
```

<br>

First, we'll just run the bare-bones `edger_multi` function.

<br>

```{r edger_multi 1, echo = T, eval = TRUE}
roi1 <- data.frame(x1 = 894, y1 = 538, x2 = 974, y2 = 1219)
edger::edger_multi(c(im1, im2), roi = list(roi1), color = "green1")
```


<br>

This function comes with a progress bar, courtesy of the `progressr` package. The progress bar tells the user the percent of images it has finished recoloring.

<br>

Now let's shift and rotate the recolored pixels. The `shift` input acts like a coordinate pair where the first value is for the x shift and the second value is for the y shift. Positive values are for shifts right or up, and negative values are for shifts left or down. The `rotate` input is the number of degrees to rotate the pixels. Let's shift the recolored pixels 200 to the right and 200 down and then rotate them 30 degrees clockwise.

<br>

```{r edger_multi 2, echo = T, eval = TRUE}
edger::edger_multi(c(im1, im2), roi = list(roi1), 
                     shift = c(200, -200), rotate = 30, color = "yellow")
```


<br>

Hopefully you can see how this functionality will be very useful if the camera viewshed changes partway through the camera's deployment. `edger` won't throw a fit if you shift or rotate any of the pixels outside of the bounds of the image; it just won't recolor them.

<br>

Now let's use two reference images. Each reference image needs its own data frame of coordinates to the ROI. Then each separate data frame gets added to a list. If one reference images needs more regions drawn on it than another, you can just set `regions` to the maximum number of regions you want in any image. We'll set `regions` to 2 and add a second row to the data frame for the first image so we can capture both snow stakes in the first image. But since we only have one row in the data frame for the second image, there will only be one ROI in the second image.

<br>

```{r edger_multi 3, echo = T, eval = TRUE}
roi1 <- data.frame(x1 = 894, y1 = 538, x2 = 974, y2 = 1219)
roi2 <- data.frame(x1 = 457, y1 = 501, x2 = 526, y2 = 1105)
roi3 <- data.frame(x1 = 633, y1 = 638, x2 = 874, y2 = 799)

edger::edger_multi(c(im1, im2), th = c(.1,.05),
                   ref_images = 2, regions = 2, 
                   roi = list(rbind(roi1, roi2),
                                 roi3), 
                   color = "deeppink1")
```


<br>

Cool! Now we've drawn both snow stakes from the first image onto the second image, and we've drawn the outlines of that bent-down tree in the second image onto the first image. This functionality is useful if you have multiple objects of interest but they don't all appear in the same image.

<br>

### Parallel Processing

<br>

Most cameras are collecting a lot more than 2 images that you'll have to recolor. That can be a lot of processing time. You can speed things up by setting the `process` argument of the `edger_multi` function to "parallel." This will call the `parallel` and `furrr` packages and allow your computer to process the images across several cores. You'll also have to set the `cores` argument to the number of cores you want it to use. It's generalkly safest to use at least one less than the total number of cores on your computer, which you can find out using `parallel::detectCores()`.

<br>
